---
title: |
  Final Project Proposal: Statistical Learning (UNL STAT 983)
subtitle: | 
  Imbalanced Classification and Prediction of Wine Quality
author:
  - name: Sarah Aurit
    affil: a
  - name: Alison Kleffner
    affil: a
  - name: Emily Robinson
    affil: a
affiliation:
  - num: a
    address: |
      Department of Statistics, University of Nebraska - Lincoln,
bibliography: references.bib
output: 
  pdf_document:
    template: template.tex
preamble:
  \usepackage{graphicx}
  \usepackage[dvipsnames]{xcolor} % colors
  \newcommand{\er}[1]{{\textcolor{blue}{#1}}}
  \newcommand{\ak}[1]{{\textcolor{RedViolet}{#1}}}
  \newcommand{\sa}[1]{{\textcolor{OliveGreen}{#1}}}
  \graphicspath{{}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	fig.align = "center"
)
library(readr)
library(tidyverse)
library(knitr)
```

<!-- \er{Emily}  -->
<!-- \ak{Alison}  -->
<!-- \sa{Sarah} -->

<!-- If you put each sentence on a new line, it makes it easier for Github to merge changes if we are all working on the paper. -->


# Introduction

Successful marking campaigns and productive selling strategies are directly linked to communication about key indicators of quality; hence, objective measurements of quality are essential. 
Within the wine industry, there are two types of quality assessment: physiochemical and sensory tests. 
Sensory tests require a human expert to assess the quality of wine based on visual, taste, and smell \citep{hu2016classification}. 
Hiring human experts to conduct sensory tests can take time and be expensive \citep{gupta2018selection}. 
In addition, taste is the least understood of all human senses \citep{cortez2009using}. 
Unlike sensory tests, laboratory tests for measuring the physiochemical characteristics of wine such as acidity and alcohol content do not require a human expert. 
The relationship between physiochemcial and sensory analysis is not well understood. 
Recently, research in the food industry has utilized statistical learning techniques to evaluate widely available characteristics of wine. This type of evaluation allows the automation of quality assessment processes by minimizing the need of human experts \citep{gupta2018selection}. 
These techniques also have the advantage of identifying important the phsiochemical characteristics that have an impact on the quality of wine as determined by a sensory test. 
In this report, we will investigate the accuracy of three classification techniques, and also address the challenges of classification with imbalanced data.

# Data

We will be working with the "Wine Quality Data Set" found on the UCI Data Repository at https://archive.ics.uci.edu/ml/datasets/Wine+Quality. \citep{UCIdataset, cortez2009using}. 
This data set consists of white (4898 samples) and red (1599 samples) Portuguese "Vinho Verde" wine samples. 
Each wine sample was processed to obtain measures for 11 physiochemical variables: fixed acidity, volatile acidity, citric acid, residual sugar, chlorides, free sulfur dioxide, total sulfur dioxide, density, pH, sulphates, and alcohol. 
A sensory test conducted by at least three sommeliers was used to determine a quality rating on an 11 point scale from 0 - very bad to 10 - very excellent.
Following \cite{hu2016classification}, we separated the wine samples in to 3 classes: Low Quality $(\le 4)$, Normal $(5-7)$, and High Quality $(\ge 8)$. 
<!-- Basic descriptives are provided below; the Mann-Whitney and chi-square tests were used for comparisons of continuous and discrete variables, respectively. -->
<!-- \begin{figure} -->
<!--   \centering -->
<!--   \includegraphics{"../results/Descriptives.png"} -->
<!-- \end{figure} -->

The quality categorization poses a challenge of working with imbalanced classes as there are many more normal wines than low or high quality wines (Figure \ref{fig:class-distribution}). 
Below, we propose resampling techniques to improve the accuracy of classification.

```{r data}
winequality <- read.csv("../data/winequality-all.csv")
```

```{r quality-distribution,include = FALSE}
winequality %>%
  mutate(quality = factor(quality)) %>%
  group_by(type, quality) %>%
  summarise(count = n()) %>%
  mutate(prop = count/sum(count)) %>%
  ungroup() %>%
  ggplot(aes(x = quality, y = prop, fill = quality)) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  facet_grid(~type) +
  theme_bw() +
  theme(aspect.ratio = 0.8) +
  scale_fill_viridis_d(end = 0.95) +
  scale_x_discrete("Quality \n (Scale 0 - 10)") +
  scale_y_continuous("Proportion", limits = c(0, 0.5), breaks = seq(0, 0.5, 0.1), labels = scales::percent)
```

```{r class-distribution, fig.height = 2.5, fig.width = 5, fig.align = 'center', fig.cap = "Wine quality class imbalance."}
winequality %>%
  mutate(qualityclass = factor(qualityclass, levels = c("Low", "Normal", "High"))) %>%
  group_by(type, qualityclass) %>%
  summarise(count = n()) %>%
  mutate(prop = count/sum(count)) %>%
  ungroup() %>%
  ggplot(aes(x = qualityclass, y = prop, fill = qualityclass, label = paste(100*round(prop, 3), "%", sep = ""))) +
  geom_bar(stat = "identity", show.legend = FALSE) +
  geom_text(vjust = -0.5, size = 3) +
  facet_grid(~type) +
  theme_bw() +
  theme(aspect.ratio = 0.6) +
  scale_fill_viridis_d(begin = 0.4, end = 0.8) +
  scale_x_discrete("Quality Class") +
  scale_y_continuous("Proportion", limits = c(0, 1.1), breaks = seq(0, 1, 0.2), labels = scales::percent)
```

<<<<<<< HEAD
```{r descriptives}
#https://cran.r-project.org/web/packages/Gmisc/vignettes/Descriptives.html;
#For labeling we use the label() function from the Hmisc package
install.packages("Hmisc")
install.packages("Gmisc")
=======
```{r descriptives, message=FALSE, warning=FALSE}
# https://cran.r-project.org/web/packages/Gmisc/vignettes/Descriptives.html;
# For labeling we use the label() function from the Hmisc package
# install.packages("Hmisc")
# install.packages("Gmisc")
>>>>>>> 8997523d6598d70c056e73498b3f7b231cda9dcb
library(Hmisc)
library(Gmisc)
data(winequality)

#Style all of the table output using htmlTable's theme handler
htmlTable::setHtmlTableTheme(css.rgroup = "")

winequality$qualityclass <- factor(winequality$qualityclass, levels=c("Low","Normal","High"),
                            labels=c("Low (N=246)","Normal (N=6,053)","High (N=198)")) 
winequality$type <- factor(winequality$type, levels=c("red","white"),
                    labels=c("Red Wine","White Wine"))

label(winequality$qualityclass)         <- "Quality"
label(winequality$type)                 <- "Type"
label(winequality$fixed.acidity)        <- "Fixed Acidity"
units(winequality$fixed.acidity)        <- "(g tartaric acid/dm^3), median (IQR)"
label(winequality$volatile.acidity)     <- "Volatile Acidity"
units(winequality$volatile.acidity)     <- "(g acetic acid/dm^3), median (IQR)"
label(winequality$citric.acid)          <- "Citric Acid"
units(winequality$citric.acid)          <- "(g/dm^3), median (IQR)"
label(winequality$residual.sugar)       <- "Residual Sugar"
units(winequality$residual.sugar)       <- "(g/dm^3), median (IQR)"
label(winequality$chlorides)            <- "Chlorides"
units(winequality$chlorides)            <- "(g sodium chloride/dm^3), median (IQR)"
label(winequality$free.sulfur.dioxide)  <- "Free Sulfur Dioxide"
units(winequality$free.sulfur.dioxide)  <- "(mg/dm^3), median (IQR)"
label(winequality$total.sulfur.dioxide) <- "Total Sulfur Dioxide"
units(winequality$total.sulfur.dioxide) <- "(mg/dm^3), median (IQR)"
label(winequality$density)              <- "Density"
units(winequality$density)              <- "(g/cm^3), median (IQR)"
label(winequality$pH)                   <- "pH"
units(winequality$pH)                   <- "median (IQR)"
label(winequality$sulphates)            <- "Sulphates"
units(winequality$sulphates)            <- "(g potassium sulphate/dm^3), median (IQR)"
label(winequality$alcohol)              <- "Alcohol"
units(winequality$alcohol)              <- "(volume %), median (IQR)"

getTable1Stats <- function(x, digits = 0, ...){
  getDescriptionStatsBy(x = x, 
                        by = winequality$qualityclass,
                        digits = digits,
                        continuous_fn = describeMedian,
                        header_count = TRUE,
                        ...)
}

t1 <- list()
t1[["Type"]] <- getTable1Stats(winequality$type)
t1[["Fixed Acidity"]] <- getTable1Stats(winequality$fixed.acidity)
t1[["Volatile Acidity"]] <- getTable1Stats(winequality$volatile.acidity)
t1[["CitricAcidity"]] <- getTable1Stats(winequality$citric.acid)
t1[["Residual Sugar"]] <- getTable1Stats(winequality$residual.sugar)
t1[["Chlorides"]] <- getTable1Stats(winequality$chlorides)
t1[["Free Sulfur Dioxide"]] <- getTable1Stats(winequality$free.sulfur.dioxide)
t1[["Total Sulfur Dioxide"]] <- getTable1Stats(winequality$total.sulfur.dioxide)
t1[["Density"]] <- getTable1Stats(winequality$density)
t1[["pH"]] <- getTable1Stats(winequality$pH)
t1[["Sulphates"]] <- getTable1Stats(winequality$sulphates)
t1[["Alcohol"]] <- getTable1Stats(winequality$alcohol)

mergeDesc(t1,
          htmlTable_args = list(caption  = "Basic Descriptive Statistics from the Wine Quality Dataset"))

t1

```
 
# Proposed Methodology

In this paper we are going to be looking at two main research questions. The first one focuses on developing a model that would work well to classify wines into three categories, which are: poor quality, normal quality and high quality. As stated in the problem description, it is desirable to classify wines using physicochemical properties since this does not involve human bias that would come into play with human tasters. We plan on using three different classification techniques. First, we will utilize random forests given prior investigation of the white wine data showed this technique seemed to work well; we would like to test this method using both white and red wine. Our second method is XGBoost, which is a gradient boosting framework that can be done in R. Previous papers on this data set have used different versions of gradient boosting like adaptive boost. Our final proposed method is KNN, as this seems to be a method that has not yet been used in any sort of wine classification. $\\$

Additionally, as described above, an issue with our data is the imbalanced nature of the data set, with most of the wines being classified into the normal quality category. We are also interested in looking at different resampling techniques in order to determine if resampling is necessary here, and to identify which method of resampling method is best. We plan on running each of the classfication methods to find predictions in conjunction with each of the resampling techniques for the sake of comparison. Initially, we will first look at the results of the classfication method predictions if no resampling is done. Since the data is very imbalanced, we would expect this to probably not do well, but it gives a baseline. The second resampling method we will test is SMOTE, which is an algorithm where the minority class (in this case low and high quality), will be oversampled. This method runs the risk of overfitting the mode. The final resampling method that we will look at is random under sampler, where majority class data are randomly removed from the data set. This risks losing valuable information in our model. We are looking for the best prediction model using the combination of resampling and classfication method, which will be determined using MSE. 

```{r Random Forest}
install.packages("randomForest")
library(randomForest)

set.seed(220355)

#Create Loop i = 1 to B

#Fit Random Forest Model with Data
model1 <- randomForest(quality ~ ., data=winequality_i) #replace with data, also move to classification
model1
which.min(model1$mse) #number of trees that produce the lowest test MSE
sqrt(model1$mse[which.min(model1$mse)]) #find RMSE of best model
plot(model1)
varImpPlot(model1) #need to remove quality variable

#Tune the model
a=c()
i=5
for (i in 3:8) {
    model1_tuned <- randomForest(quality ~ ., data=winequality_i, ntree=500, mtry=i,importance=TRUE) 
    predtune <- predict(model1_tuned, winequality_i, type="class") 
    a[i-2] = mean(predtune == winequality_i$quality) 
}
a
plot(3:8,a)

#Fit Random Forest Model 
predTest <- predict(model1_tuned, winequality_i, type="class") 
mean(predTest ==winequality_i$qualityclass) 
table(predTest, winequality_i$qualityclass) 
importance(model1_tuned)
varImpPlot(model1_tuned)

```
 
